---
title: "Thesis_Airborne_Transmission"
output: html_notebook
---

Install packages and open libraries as needed
```{r Install packages and libraries needed}
# install.packages("tidyverse")
# install.packages("RCurl") # Pull data from Github Raw version 
# install.packages("sensitivity")
# install.packages("ggplot2")
# install.packages("readxl")
# install.packages("reshape2")
library(readxl)
library(deSolve)
library(data.table)
library(data.table)
library(ODEsensitivity)# If you want to use the s
library(ggplot2)
library(plotly)
library(lattice)#Making heat maps
library(latticeExtra) #make Heat maps
library(ggplot2)
library (RCurl)
library (reshape2)
```

Insert Data for parameters. Most can be used as a guide on what to use for different scenarios. Importing CSV files but see excel files for sources (https://github.com/copelaa/Airborne_Droplet_Transmission_CopelandThesis/tree/master/Data).

CSV files hosted on github and can be manually downloaded and opened from local machine, or pulled directly from Github. 
```{r Input Data Tables From GitHub}
##Risk per distance (R)if D=x
Risk_factor = "https://raw.githubusercontent.com/copelaa/Airborne_Droplet_Transmission_CopelandThesis/master/Data/Room%20Info.csv"

Risk_factor <- read.csv(Risk_factor)

##Quanta of virus (q)
Virus_Quanta = "https://raw.githubusercontent.com/copelaa/Airborne_Droplet_Transmission_CopelandThesis/master/Data/Virus%20Quanta.csv"

Virus_Quanta=read.csv(Virus_Quanta)

##Pulmonary Rate
Pulmonary_Rate = "https://raw.githubusercontent.com/copelaa/Airborne_Droplet_Transmission_CopelandThesis/master/Data/Pulmonary%20rate.csv"

Pulmonary_Rate = read.csv(Pulmonary_Rate)
Pulmonary_Rate$p_gendermean <- rowMeans(Pulmonary_Rate[,c('Female..m3.hr.', 'Male..m3.hr.')], na.rm=TRUE)

##ACH changes based on Volume and CFM per occupancy type
Ventilation = "https://raw.githubusercontent.com/copelaa/Airborne_Droplet_Transmission_CopelandThesis/master/Data/Ventilation.csv"
Ventilation = read.csv(Ventilation)
```


Build out scenarios you want to Test or compare below I will start with building out a typical **Office** Scenario. Where:
- Typical ventilation rates are used that meet American ASHRAE Code 
- Pulmonary rates for typical working person 21 years old - 61 years old. 
- Quanta is currently estimated for SARS COVID 2 at 100. Other more known quanta are provided with infectious levels varying. 
- Risk factor can be chosen as one point but depending on event we will come up with an estimate later on in chunk {r Risk Factor Estimation}. 
```{r Office Scenerio}

##OFFICE

#Risk set below is is all people were the same distance from the infected source 
Risk <- subset.data.frame(Risk_factor, D ==1.0) 
rownames(Risk)=NULL
Risk <- tibble::column_to_rownames(Risk, "D")
Risk <- Risk$R

#Same as above but using distance 
# distance <- subset.data.frame(Risk_factor, D ==1.0) 
# distance <-tibble::column_to_rownames(distance, "R")
# distance <- distance$D
# distance <-as.numeric(distance)

##Quanta (q)
q <- 178

##Pulmonary rate (p)
p <- subset.data.frame(Pulmonary_Rate, Age == "21-61" & Activity == "SEDENTARY", select = "p_gendermean")
p <- p$p_gendermean
p <-as.numeric(p)

##ACH (a) 
af <- 4 #Air Changes Per Hour
af <- as.numeric(af)

##Volume (V)
Vf <- 226.53 # in meters - Open Office space 1000 sq ft(93m) 8ft (2.43m) ceiling height 
Vf <- as.numeric(Vf)

beta_office <-(Risk +((p*q)/(Vf*af)))           # Transmission risk at office space

```

Below is the Airplane scenarios parameters defined: 
- Typical ventilation rates are used that meet American ASHRAE Code 
- Pulmonary rates for typical working person 21 years old - 61 years old. 
- Quanta is currently estimated for SARS COVID 2 at 100. Other more known quanta are provided with infectious levels varying. 
- Risk factor can be chosen as one point but depending on event we will come up with an estimate later on in chunk 'Risk Factor Estimations. **Note** Distance of susceptible persons will be limited to normal Boeing 777-300ER ventilation zones (typically 7 zones in main cabin )
```{r Plane Parameters}

##PLANE
Risk_plane<-.013 # Again using a static distance of 2.0 M or 2.5 M away from infected source

##Quanta (q)
#Same as Office 

##Pulmonary rate (p)
#Same as Office 

##ACH (a)
ap <- 15 # Based on suggested code 15-20 is standard
ap <- as.numeric(ap)

##Volume (V)# ~ 13ft (31 inches per row x 5 rows) x 19 ft. 3 in (Cabin width) x 73In Height Estimated 42.851 meter3 or 1,522ft3
Vp<- 3.937*5.87*1.8542 #in meters
```


```{r Additional Parameters}
          r= 1/6 #Recover rate (gamma)
          g= 1/4 #Transition rate from latent to infectious - 4 day (4/365 year) latent period 
          f= (2.2/100)#Case fatality ratio

            
# Establish Time
dt_office <- seq(0.01,1,.01) #This gives start and end times (0 to 1 day) and the time step 14.4mins (100 measurements within 24 hours) 
dt_plane <- seq(0.005,1,.005) #This gives start and end times (0 to 1 day) and the time step 7.2 mins
```

As mentioned previously it is not realistic that all susceptible people would be the same distance away from the infected source unless we were only looking at a small subset of persons within an enclosed space. 

In order to find a more realistic risk rate, a random points that correspond with the number of susceptible people will be plotted over a grid of corresponding to the spaces dimensions (height and width)

First Create a function to calculate Risk Based on Distance from Index source:
```{r Calculate Risk based as a function of Distance}
calcRisk <- function(distance){
  if (distance < 1){
    risk <- 0.13
  } else if (distance < 2) {
    risk <- 0.026}
    else {risk <- 0.026/2^(as.integer(distance-1))}
  return(risk)
}
```


Random Cumulative Risk in Office
```{r}
#Following 4 lines are if you want to get the average distance of 1000 iterations and then the average risk 

#iterations <- 10000
#mean_distances_office <-rep(0, iterations)
#for(i in 1:iterations){
#set.seed(i)

set.seed(123)
{
  min_x <- 0
  max_x_office <- 10
  min_y <- 0
  max_y_office<- 10
  x_step <- 0.3
  y_step <- 0.3
  x_range <- seq(min_x, max_x_office, x_step)
  y_range <- seq(min_y, max_y_office, y_step)
  num_points_office <- 20 #population enclosed space 
  xs = sample(x_range, num_points_office, replace = TRUE, prob = NULL)
  ys = sample(y_range, num_points_office, replace = TRUE, prob = NULL)
  
  #Randomly plot points in office space 
  sample_points_office <- array(c(xs, ys), dim = c(num_points_office, 2))
  
Distance_infected_source_office <- as.data.frame(sweep(sample_points_office[1:20,1:2],2,sample_points_office[1,1:2]))#Finding points distance from Random infected source location (1st point indicated in sample_point_office dataframe)

mean_distance_office = sqrt((Distance_infected_source_office$V1)^2 + (Distance_infected_source_office$V2)^2)#Find the distance of each point from infected source 

Distance_infected_source_office$distance_from_index <- mean_distance_office #Create new Dataframe of just distance from Index case

#Drop Infected source
Distance_infected_source_office<- Distance_infected_source_office[-c(1),] 
#If you want to find mean distance and then risk of 1000 random iterations use the below line

#mean_distances_office[i] <- mean(mean_distance_office)}
}

#Graph random point in enclosed space
ggplot(as.data.frame(sample_points_office), aes(x=xs, y=ys)) +
  
  geom_point(size=2, shape=23, col='red') #Plot Location of People

#Find Risk associated with Distance
Cumulative_risk_office<- as.numeric(lapply(Distance_infected_source_office[ ,3], calcRisk))

#Add Risk to Dataframe
Distance_infected_source_office$Risk <- Cumulative_risk_office
#Find cumulative risk and distance
Cumulative_risk_office<- sum(Cumulative_risk_office)/19 #people in enclosed space minus infected sources
Cumulative_distance_office<- mean(Distance_infected_source_office$distance_from_index)

```

Now that we have all the parameters defined we can create the contour plots to show how Air Changes and distance may affect transmission

```{r OFFICE Contour Plot - Beta}
##Risks Beta for all controllable measures 
## Setting range for contour plot 
Risk_fac_Cont_office <- Risk_factor[1:10,2]
ach_office <- c(1:10)
Risk_fac_dist_office <- Risk_fac_dist_office <- as.numeric(rep(seq(from = .5, to = 5,by = .5), each=10))
 

#Expand Dataframe to include all risk and distance at each ACH level (1-10)
RiskVa_office<-expand.grid(ach_office = seq(1, 10, 1), Risk_fac_Cont_office = Risk_fac_Cont_office)

##Beta using all risks
n_office <- 100
beta_office <- rep(1, n_office)
for(i in 1:100) {
  beta_office[i] <- (RiskVa_office[i,2] + ((p*q)/(Vf*RiskVa_office[i,1])))
}
print(beta_office)

RiskVa_office$beta <- beta_office

#Add distance 
RiskVa_office$Risk_fac_dist_office<-Risk_fac_dist_office

# #Reorder BASED  on Distance
RiskVa_office$Risk_fac_dist_office <- as.numeric(as.character(RiskVa_office$Risk_fac_dist_office))
RiskVa_office<- RiskVa_office[order(RiskVa_office$Risk_fac_dist_office),]


## Level Plot
x.labels <- seq(1,5,.5)

 obj_office <- if (require("mgcv"))
  levelplot(beta_office ~ Risk_fac_Cont_office * ach_office, RiskVa_office, panel = panel.2dsmoother,
            form = z ~ s(x, y), method = "gam", n=500,
            cuts = 10, contour=TRUE,
            xlab = "Distance from index case (m)",
            ylab = "Air Changes (m3/h)",
            xlim = c(.13, 0.00325),xscale.components.default = x.labels,
            ylim = c(1, 10),
            main = "Beta Parameter", sub="Office")


update(obj_office, par.settings =
         custom.theme(region = colorspace::diverge_hsv(9),
                      bg = "grey90", fg = "grey20", pch = 10))
```

Change Beta to R0 in Contour plot 
```{r OFFICE Contour Plot - R0}

## plot changing R0 under control (R0)Beta/gamma
R0_office <- beta_office/(r)
matplot(dt_office,R0_office,type='l',ylab='R0',xlab='Time (hours)',lty=1)
abline(h=1, col="red")


###Expand R0/   beta / time of event () - (at random distribution )
objR0_office <- if (require("mgcv"))
  levelplot(R0_office ~ Risk_fac_Cont_office*ach_office, RiskVa_office, panel = panel.2dsmoother,
            form = z ~ s(x, y), method = "gam", n=500,
            cuts = 10, contour=TRUE,
            xlab = "Distance from index case(m)",
            ylab = "Air Changes m3/h",
            xlim = c(min(.13),max(0.00325)),xscale.components.default = x.labels, ylim = c(min(ach_office), max(ach_office)),
            main = "R0 related to ACH and Distance", sub="Office")
update(objR0_office, par.settings =
         custom.theme(region = colorspace::diverge_hsv(9),
                      bg = "white", fg = "grey20", pch = 10))
```


If you want to run a compartment model for the OFFICE: 
```{r Office - Compartment model SEIR}
#establish additional parameters 
params_office<- c(bf=(Cumulative_risk_office +((p*q)/(Vf*af))), #Transmission probability
                  r= 1/6, #Recover rate (gamma)
                  g= 1/4, #Transition rate from latent to infectious - 4 day (4/365 year) latent period 
                  f= (2.2/100)#Case fatality ratio
)
inits_office <- c(S=19,E=0, I=1, R=0) #This line gives the initial values each compartment
SEIR_office <- function(t, x, params_office){ 
  
  with(as.list(c(params_office,x)),{
    
    
    #Flow in and out of each compartment
    dS <- -bf*S*I/(S+E+I+R) 
    dE <- (bf*S*I/(S+E+I+R)) - (g*E)
    dI <- (g*E) - (r*I) 
    dR <- r*I 
    
    
    #compartments
    der_office <- c(dS,dE,dI,dR)
    list(der_office) # the output must be returned
  }) 
}

#lsoda to formats the output as a data frame
data_out_office <- as.data.frame(lsoda(inits_office, dt_office, SEIR_office, parms=params_office))

#This line plots all of your compartments. 
#The syntax for matplot is (x value, y value, graphical parameters)
#Here, x is the first column of data_out (time), y is each of columns 2 through 5 (S, E, I, and R), and saying to plot solid lines (type="l", lty=1) 
#By default, each line plotted will be given a sequential color (S will be color 1, E color 2, etc.)
matplot(data_out_office[,1],data_out_office[,2:5],type='l',ylab='People',xlab='Time (day)',lty=1)

```

Find Cumulative Risk of persons in vicinity of infected source 
```{r Risk Factor Estimation - Plane}
#Following 4 lines are if you want to get the average distance of 1000 iterations and then the average risk 

#iterations <- 10000
#mean_distances_plane <-rep(0, iterations)
#for(i in 1:iterations){
#set.seed(i)

set.seed(123)
{
  min_x <- 0
  max_x_plane<- 4# Length in M
  min_y <- 0
  max_y_plane<- 6 #Width in M
  x_step <- 0.3
  y_step <- 0.3
  x_range <- seq(min_x, max_x_plane, x_step)
  y_range <- seq(min_y, max_y_plane, y_step)
  num_points_plane <- 10 #population enclosed space 
  xs = sample(x_range, num_points_plane, replace = TRUE, prob = NULL)
  ys = sample(y_range, num_points_plane, replace = TRUE, prob = NULL)
  
  #Randomly plot points in office space 
  sample_points_plane <- array(c(xs, ys), dim = c(num_points_plane, 2))
  
Distance_infected_source_plane <- as.data.frame(sweep(sample_points_plane[1:10,1:2],2,sample_points_plane[1,1:2]))#Finding points distance from Random infected source location (1st point indicated in sample_point_office dataframe)

mean_distance_plane = sqrt((Distance_infected_source_plane$V1)^2 + (Distance_infected_source_plane$V2)^2)#Find the distance of each point from infected source 

Distance_infected_source_plane$distance_from_index <- mean_distance_plane #Create new Dataframe of just distance from Index case

#Drop Infected source
Distance_infected_source_plane<- Distance_infected_source_plane[-c(1),] 
#If you want to find mean distance and then risk of 1000 random iterations use the below line

#mean_distances_plane[i] <- mean(mean_distance_plane)}
}

#Graph random point in enclosed space
ggplot(as.data.frame(sample_points_plane), aes(x=xs, y=ys)) +
  
  geom_point(size=2, shape=23, col='red') #Plot Location of People

#Find Risk associated with Distance
Cumulative_risk_plane<- as.numeric(lapply(Distance_infected_source_plane[ ,3], calcRisk))

#Add Risk to Dataframe
Distance_infected_source_plane$Risk <- Cumulative_risk_plane
#Find cumulative risk and distance
Cumulative_risk_plane<- sum(Cumulative_risk_plane)/9 #people in enclosed space minus infected sources
Cumulative_distance_plane<- mean(Distance_infected_source_plane$distance_from_index)
```

Now that we have the parameters indicated for the scenarios we want to test we will find the preliminary beta coefficients 
```{r Plane Contour Plot - Beta}
##Risks Beta for all controllable measures 
## Setting range for contour plot 
Risk_fac_Cont_plane <- Risk_factor[1:10,2]
ach_plane <- c(1:20)
Risk_fac_dist_plane <- as.numeric(rep(seq(from = .5, to = 5,by = .5), each=20))


#Expand Dataframe to include all risk and distance at each ACH level (1-10)
RiskVa_plane<-expand.grid(ach_plane = seq(1, 20, 1), Risk_fac_Cont_plane = Risk_fac_Cont_plane)

##Beta using all risks
n_plane <- 200
beta_plane <- rep(1, n_plane)
for(i in 1:200) {
  beta_plane[i] <- (RiskVa_plane[i,2] + ((p*q)/(Vf*RiskVa_plane[i,1])))
}
print(beta_plane)

RiskVa_plane$beta <- beta_plane

#Add distance 
RiskVa_plane$Risk_fac_dist_plane<-Risk_fac_dist_plane

# #Reorder BASED  on distance
RiskVa_plane$Risk_fac_dist_plane <- as.numeric(as.character(RiskVa_plane$Risk_fac_dist_plane))
RiskVa_plane<- RiskVa_plane[order(RiskVa_plane$Risk_fac_dist_plane),]


## Level Plot
x.labels <- seq(1,5,.5)

 obj_plane <- if (require("mgcv"))
  levelplot(beta_plane ~ Risk_fac_Cont_plane * ach_plane, RiskVa_plane, panel = panel.2dsmoother,
            form = z ~ s(x, y), method = "gam", n=500,
            cuts = 10, contour=TRUE,
            xlab = "Distance from index case (m)",
            ylab = "Air Changes (m3/h)",
            xlim = c(.13, 0.00325),xscale.components.default = x.labels,
            ylim = c(1, 20),
            main = "Beta Parameter", sub="Plane")


update(obj_plane, par.settings =
         custom.theme(region = colorspace::diverge_hsv(9),
                      bg = "grey90", fg = "grey20", pch = 10))
```
Change Beta to R0 in Contour plot 

```{r Plane Contour Plot - R0}
## plot changing R0 under control (R0)Beta/gamma
R0_plane <- beta_plane/(r)
matplot(dt_plane,R0_plane,type='l',ylab='R0',xlab='Time (hours)',lty=1)
abline(h=1, col="red")


###Expand R0/   beta / time of event () - (at random distribution )
objR0_plane <- if (require("mgcv"))
  levelplot(R0_plane ~ Risk_fac_Cont_plane*ach_plane, RiskVa_plane, panel = panel.2dsmoother,
            form = z ~ s(x, y), method = "gam", n=500,
            cuts = 10, contour=TRUE,
            xlab = "Distance from index case(m)",
            ylab = "Air Changes m3/h",
            xlim = c(min(.13),max(0.00325)),xscale.components.default = x.labels, ylim = c(min(ach_plane), max(ach_plane)),
            main = "R0 related to ACH and Distance", sub="Plane")
update(objR0_plane, par.settings =
         custom.theme(region = colorspace::diverge_hsv(9),
                      bg = "white", fg = "grey20", pch = 10))
```

If you want to create compartment plots and data based on the altered Betas:
```{r Plane Compartment Model SEIR}
#Additional Parameters 

params_plane<- c(bp=(Cumulative_risk_plane +((p*q)/(Vp*ap))), #Transmission probability
                  r= 1/6, #Recover rate (gamma)
                  g= 1/4, #Transition rate from latent to infectious - 4 day (4/365 year) latent period 
                  f= (2.2/100)#Case fatality ratio
)
inits_plane <- c(S=9,E=0, I=1, R=0) #This line gives the initial values each compartment
SEIR_plane <- function(t, x, params_plane){ 
  
  with(as.list(c(params_plane,x)),{
    
    
    #Flow in and out of each compartment
    dS <- -bp*S*I/(S+E+I+R) 
    dE <- (bp*S*I/(S+E+I+R)) - (g*E)
    dI <- (g*E) - (r*I) 
    dR <- r*I 
    
    
    #compartments
    der_plane <- c(dS,dE,dI,dR)
    list(der_plane) # the output must be returned
  }) 
}

#lsoda to formats the output as a data frame
data_out_plane <- as.data.frame(lsoda(inits_plane, dt_plane, SEIR_plane, parms=params_plane))

#This line plots all of your compartments. 
#The syntax for matplot is (x value, y value, graphical parameters)
#Here, x is the first column of data_out (time), y is each of columns 2 through 5 (S, E, I, and R), and saying to plot solid lines (type="l", lty=1) 
#By default, each line plotted will be given a sequential color (S will be color 1, E color 2, etc.)
matplot(data_out_plane[,1],data_out_plane[,2:5],type='l',ylab='People',xlab='Time (day)',lty=1)
```

