---
title: "Thesis_Airborne_Transmission"
output: html_notebook
---

Install packages and open libraries as needed
```{r Install packages and libraries needed}
# install.packages("tidyverse")
# install.packages("RCurl") # Pull data from Github Raw version 
# install.packages("sensitivity")
# install.packages("ggplot2")
# install.packages("readxl")
library(readxl)
library(deSolve)
library(data.table)
library(data.table)
library(ODEsensitivity)# If you want to use the s
library(ggplot2)
library(plotly)
library(lattice)#Making heat maps
library(latticeExtra) #make Heat maps
library(ggplot2)
library (RCurl)
```

Insert Data for parameters. Most can be used as a guide on what to use for different scenarios. Importing CSV files but see excel files for sources (https://github.com/copelaa/Airborne_Droplet_Transmission_CopelandThesis/tree/master/Data).

CSV files hosted on github and can be manually downloaded and opened from local machine, or pulled directly from Github. 
```{r Input Data Tables From GitHub}
##Risk per distance (R)if D=x
Risk_factor = "https://raw.githubusercontent.com/copelaa/Airborne_Droplet_Transmission_CopelandThesis/master/Data/Room%20Info.csv"

Risk_factor <- read.csv(Risk_factor)

##Quanta of virus (q)
Virus_Quanta = "https://raw.githubusercontent.com/copelaa/Airborne_Droplet_Transmission_CopelandThesis/master/Data/Virus%20Quanta.csv"

Virus_Quanta=read.csv(Virus_Quanta)

##Pulmonary Rate
Pulmonary_Rate = "https://raw.githubusercontent.com/copelaa/Airborne_Droplet_Transmission_CopelandThesis/master/Data/Pulmonary%20rate.csv"

Pulmonary_Rate = read.csv(Pulmonary_Rate)
Pulmonary_Rate$p_gendermean <- rowMeans(Pulmonary_Rate[,c('Female..m3.hr.', 'Male..m3.hr.')], na.rm=TRUE)

##ACH changes based on Volume and CFM per occupancy type
Ventilation = "https://raw.githubusercontent.com/copelaa/Airborne_Droplet_Transmission_CopelandThesis/master/Data/Ventilation.csv"
Ventilation = read.csv(Ventilation)
```


Build out scenarios you want to Test or compare below I will start with building out a typical **Office** Scenario. Where:
- Typical ventilation rates are used that meet American ASHRAE Code 
- Pulmonary rates for typical working person 21 years old - 61 years old. 
- Quanta is currently estimated for SARS COVID 2 at 100. Other more known quanta are provided with infectious levels varying. 
- Risk factor can be chosen as one point but depending on event we will come up with an estimate later on in chunk {r Risk Factor Estimation}. 
```{r Office Scenerio}

##OFFICE

#Risk set below is is all people were the same distance from the infected source 
Risk <- subset.data.frame(Risk_factor, D ==1.0) 
Risk <-tibble::column_to_rownames(Risk, "D")
Risk <- Risk$R

#Same as above but using distance 
# distance <- subset.data.frame(Risk_factor, D ==1.0) 
# distance <-tibble::column_to_rownames(distance, "R")
# distance <- distance$D
# distance <-as.numeric(distance)

##Quanta (q)
q <- 100

##Pulmonary rate (p)
p <- subset.data.frame(Pulmonary_Rate, Age == "21-61" & Activity == "SEDENTARY", select = "p_gendermean")
p <- p$p_gendermean
p <-as.numeric(p)

##ACH (a) 
af <- 4 #Air Changes Per Hour
af <- as.numeric(as)

##Volume (V)
Vf <- 226.53 # in meters - Open Office space 1000 sq ft 8ft ceiling height 
Vf <- as.numeric(Vs)

beta_office <-(Risk +((p*q)/(Vf*af)))           # Transmission risk at office space

```

Below is the Airplane scenarios parameters defined: 
- Typical ventilation rates are used that meet American ASHRAE Code 
- Pulmonary rates for typical working person 21 years old - 61 years old. 
- Quanta is currently estimated for SARS COVID 2 at 100. Other more known quanta are provided with infectious levels varying. 
- Risk factor can be chosen as one point but depending on event we will come up with an estimate later on in chunk 'Risk Factor Estimations. **Note** Distance of susceptible persons will be limited to normal Boeing 777-300ER ventilation zones (typically 7 zones in main cabin )
```{r Plane Parameters}

###Plane 
Risk_plane<-.013 # Again using a static distance of 2.0 M or 2.5 M away from infected source

##Quanta (q)
#Same as Office 

##Pulmonary rate (p)
#Same as Office 

##ACH (a)
ap <- 15 # Based on suggested code 15-20 is standard
ap <- as.numeric(ap)

##Volume (V)# ~ 13ft (31 inches per row x 5 rows) x 19 ft. 3 in (Cabin width) x 73In Height Estimated 42.851 meter3 or 1,522ft3
Vp<- 3.937*5.87*1.8542 #in meters
```

As mentioned previously it is not realistic that all susceptible people would be the same distance away from the infected source unless we were only looking at a small subset of persons within an enclosed space. 

In order to find a more realistic risk rate, a random points that correspond with the number of susceptible people will be plotted over a grid of corresponding to the spaces dimensions (height and width)

```{r Risk Factor Estimation}

```

Now that we have the parameters indicated for the scenarios we want to test we will find the preliminary beta coefficients 
```{r}

```

